`skater`: An R package for SNP-based Kinship Analysis, Testing, and Evaluation

### 2021-09-07 paper/paper.Rmd
change to JOSS format

---
# title_full: "skater: An R package for SNP-based Kinship Analysis and Benchmarking"
title_full: "skater: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"
title_short: "skater package for kinship analysis"

# One of the following below. See https://academic.oup.com/bioinformatics/pages/instructions_for_authors#Types%20of%20Manuscript
# one of `Original Paper`, `Discovery Note`, `Applications Note`, `Review`, `Letter to the Editor`, `Conference Paper`    
manuscript_type: Applications Note

# One of the following below. See https://academic.oup.com/bioinformatics/pages/instructions_for_authors#Detailed%20Scope%20Guidelines
# Genome analysis
# Sequence analysis
# Phylogenetics
# Structural bioinformatics
# Gene expression
# Genetic and population analysis
# Systems biology
# Data and text mining
# Databases and ontologies
# Bioimage informatics
subject_section:  Genetics and Population Analysis

access_date:
  - day: DD
    month: MM
    year: YY

author:
  - name: Stephen D. Turner
    affiliation: "1"
  - name: V.P. Nagraj
    affiliation: "1"
  - name: Matthew Scholz
    affiliation: "1"
  - name: Shakeel Jessa
    affiliation: "1"
  - name: Carlos Acevedo
    affiliation: "1"
  - name: Jianye Ge
    affiliation: "2,3"
  - name: August E. Woerner
    affiliation: "2,3"
  - name: Bruce Budowle
    affiliation: "2,3"

address:
  - code: 1
    address: Signature Science, LLC., Austin, TX 78759, USA.
  - code: 2
    address: Center for Human Identification, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
  - code: 3
    address: Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA

# Full email address to be given, preferably an institution email address.  
corresponding_author_text: "To whom correspondence should be addressed. E-mail: sturner\\@signaturescience.com"
corresponding_author_email: "sturner\\@signaturescience.com"

motivation: |
  SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation. A dedicated software package that consistently and intuitively implements this analysis functionality is needed.

results: |
  Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**. The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.

availability: |
  The skater package is implemented as an R package and is released under the MIT license at https\:\/\/github.com/signaturescience/skater. Documentation is available at https\:\/\/signaturescience.github.io/skater.

supplementary_info: |
  Supplementary data are available at Bioinformatics Online.

output:
  rticles::bioinformatics_article:
    citation_package: natbib

bibliography: bibliography.bib
---

@@

@@

<!-- The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis. Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data. The package is designed to adhere to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem [@Wickham2019]. -->

The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis. Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data. The package is designed to adhere to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem.

@@

# Funding {.unnumbered}

@@

>>>>

---
title: "`skater`: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"

tags:
  - R
  - genetics
  - kinship analysis
  - SNPs
  - sequencing

citation_author: Turner, S.D. et al.

authors:
  - name: Stephen D. Turner
    orcid: 0000-0001-9140-9028
    affiliation: "1"
  - name: V.P. Nagraj
    orcid: 0000-0003-0060-566X
    affiliation: "1"
  - name: Matthew Scholz
    orcid: 0000-0003-3686-1227
    affiliation: "1"
  - name: Shakeel Jessa
    affiliation: "1"
  - name: Carlos Acevedo
    affiliation: "1"
  - name: Jianye Ge
    orcid: 0000-0001-8724-075X
    affiliation: "2,3"
  - name: August E. Woerner
    orcid: 0000-0002-9372-1127
    affiliation: "2,3"
  - name: Bruce Budowle
    orcid: 0000-0003-4116-2930
    affiliation: "2,3"

affiliations:
  - name: Signature Science, LLC., Austin, TX 78759, USA.
    index: 1
  - name: Center for Human Identification, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
    index: 2
  - name: Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
    index: 3

date: 07 September 2021
year: 2021
bibliography: bibliography.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS

---

@@

# Summary

**Motivation:** SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation. A dedicated software package that consistently and intuitively implements this analysis functionality is needed.  
**Results:**  Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**. The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.  
**Availability:** The skater package is implemented as an R package and is released under the MIT license at https\:\/\/github.com/signaturescience/skater. Documentation is available at https\:\/\/signaturescience.github.io/skater.

@@

The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis. Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data. The package is designed to adhere to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem [@Wickham2019].

@@

# Acknowledgements

@@

# References
